---
title: "ZoningData"
author: "Tucker Schulz"
date: "2024-02-15"
output: html_document
---

```{r setup, include=FALSE}
#Knit settings 
knitr::opts_chunk$set(echo = TRUE,
                      #warning = F,
                      #message = F,
                      fig.align = "center")

# Install necessary packages
pacman::p_load(tidyverse, janitor, lubridate, ggthemes)



```

## VT Zoning Info

This is R markdown script is....

## The Data

```{r jurisdictions}
districts <- read.csv("https://github.com/thebestyoshibird/VT_Zoning_Atlas/blob/main/ZoningAtlas/consolidated/district_zoning_data.csv?raw=true") |> 
  clean_names()
```

## Filtering

To represent current completed jurisdictions, those in progress are filtered out.

```{r jurisdictionComplete, echo=TRUE}
districtsComplete <- 
  districts |> 
    #Filter to only keep completed residential districts
    dplyr::filter(status == "Completed" & type_of_zoning_district != "Nonresidential") |> 
      #Select desired columns
      dplyr::select(county, full_district_name, type_of_zoning_district:x4_family_treatment) |> 
          pivot_longer(
            cols = c(x1_family_treatment:x4_family_treatment),
            names_to = "family_treatment",
            values_to = "allowed"
          ) |> 
    dplyr::mutate(
       family_size = parse_number(family_treatment)
    )


```

## County Summaries

-   affordable_housing: a count of affordable housing districts by county 
-   multi_family_house: a count of districts with multi-family (3+) housing allowed
-   elderly_housing: a count of elderly housing districts by county 

```{r county summaries}
county_summaries <- districtsComplete |> 
  #grouped by county for county comparisons
  dplyr::group_by(county) |>
  
  dplyr::mutate(county = tolower(str_extract(county, "\\b\\w+\\b"))) |> 
  
  #creating above summary columns and removing na values
  dplyr::summarise(affordable_housing = sum(affordable_housing_district == "Yes", na.rm = TRUE),
                   multi_family_housing = sum(family_size %in% c(3, 4, 5) & 
                                               allowed %in% c("Allowed/Conditional", "Public Hearing"), na.rm = TRUE),
                   eldery_housing = sum(elderly_housing_district == "Yes", na.rm = TRUE),
                   )|> 
  dplyr::arrange(county) 
  
tibble(county_summaries)
```
## Implementing Urbnmapr to get Vermont Counties Data Frame
```{r}
vermont_counties <- map_data("county") |> 
  dplyr::filter(region == "vermont")
```

```{r}
#merging counties map with vermont counties
counties_map <- 
  left_join(x = vermont_counties,
            y = county_summaries,
            by = c("subregion" = "county")) 


```



## Graphing the Map
```{r}
ggplot() +
  geom_polygon(data = counties_map,
               mapping = aes(x = long,
                             y = lat,
                             group = group, 
                             fill = affordable_housing),
               color = "black") +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme_map() +
  theme(legend.position = "right")
```



## Multi-family housing by county

As of right now this only looks at family sizes of 3 and 4, need to make a x5 treatment column of some kind to include in the pivot longer function above.

```{r county comparisons}

countySumsGG <- districtsComplete |> 
  dplyr::filter(family_size > 2)

  



ggplot(data = countySumsGG,
       mapping = aes(x = county)) +
  
  geom_bar(aes(fill = allowed),
           position = "dodge2") + 
  labs(x = "Counties",
       y = "Multi Family Housing Districts",
       fill = "Status") +
  
  theme_bw() +
  
  scale_fill_manual(
    values = c(
      "Allowed/Conditional" = "blue",
      "Overlay" = "green",
      "Prohibited" = "red",
      "Public Hearing" = "purple"
    )
  ) +
  
  # Add this to your graph to center the title and rotate the x-axis text and remove #
   theme(
     axis.title.x = element_blank(),
     panel.background = element_blank(),
     plot.title = element_text(hjust = 0.5),
     axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
   ) +
  scale_y_continuous(expand = c(0, 0, 0.05, 0)) 

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
